
% COMMANDS: useful commands for formatting this particular document
%
\newcommand{\CranMirror}[0]{
  \href{http://cran.opensourceresources.org/}{http://cran.opensourceresources.org/}
}

\newcommand{\InstallInstructionsHref}[0]{
  \href{http://cran.r-project.org/doc/manuals/R-admin.html}{http://cran.r-project.org/doc/manuals/R-admin.html}
}

\newcommand{\MacInstallHref}[0]{
\href{http://cran.r-project.org/doc/manuals/R-admin.html\#Installing-R-under-_0028Mac_0029-OS-X}{Mac}
}
\newcommand{\WindowsInstallHref}[0]{
\href{http://cran.r-project.org/doc/manuals/R-admin.html\#Installing-R-under-Windows}{Windows}
}
\newcommand{\UnixInstallHref}[0]{
\href{http://cran.r-project.org/doc/manuals/R-admin.html\#Installing-R-under-Unix_002dalikes}{Unix-alike}
}



% Introduction Installation
\section{Introduction}

Zelig's installation procedure is straightforward, though the package itself is not standalone, and requires the installation of R version 2.13 (or greater). That is, because Zelig is written in the R statistical programming language, it cannot be installed without support for the R programming language. As a result of this, installing Zelig and Zelig-compliant packages can be divided into three tasks:

\begin{enumerate}
	\item Download and Install R,
	\item Install Zelig, and
	\item Install Optional Zelig Add-ons
\end{enumerate}

The following guide is intended to quickly and easily explain each of these steps.

% Requirements for Installation: list the two requirements
%
%
\section{Requirements for Installation}

The Zelig software suite has only two requirements: 

\begin{enumerate}
	\item {\bf R version 2.13+}, which can be downloaded at \href{http://www.r-project.org/}{\tt http://r-project.org/}
	\item A major operating system, either:
		\begin{itemize}
			\item Mac OS X 10.4+,
			\item Windows or
			\item Linux
		\end{itemize}
\end{enumerate}

Installation instructions for R can be found on the R-project website. Simply visit the download page, and select any mirror link, though this one is recommended:

\CranMirror



% Installing R
%
%
\section{Installing R}

Installing R is typically straightforward, regardless of which operating system is being used. Several useful documents exist on CRAN (The Comprehensive R Archive Network) for explanation and troubleshooting of R installation. These documents can be found on any CRAN mirror. Specifically, the complete guide to install R can be found at:

\InstallInstructionsHref

This document contains specific documents for installing R on \MacInstallHref, \WindowsInstallHref, and \UnixInstallHref systems.

% Easy Installation Instructions: explain how to install Zelig via install.R
%
%
\section{Easy Installation}

Once R version 2.13 or greater has been installed on the client's machine, setting up Zelig is a breeze. R has built-in facilities for managing the installation of statistical packages. This provides a simple mechanism for installing Zelig, regardless of the operating system that is being used.

To install Zelig, as well as its add-on packages, simply:

\begin{enumerate}
	\item {\bf Install R version 2.13 or greater}. Download R from the R project's website, which can be found at \CranMirror
  \item {\bf Launch R}. Once R is installed, this program can be found wherever the computer stores applications (e.g. ``Program Files" on a Windows machine)
  \item At the R command-prompt, type:
    \begin{verbatim}
    source("http://gking.harvard.edu/zelig/install.R")
    \end{verbatim}
    This launches pre-written install script, which directs R to download all the appropriate statistical packages associated with the Zelig software suite.
\end{enumerate}



% Advanced Installation
%
%
\section{Advanced Installation}

For users familiar with R and Zelig, it may be useful to selectively install packages. In order to do this, users simply need to use the {\tt install.packages} function built into R's functionality.

\subsection{Install Zelig without Additional Packages}
This installation procedure will install Zelig without any add-on packages. That is, Zelig will only download files necessarily for developing \emph{new} Zelig packages and basic generalized linear model regressions - logit, gamma, gaussian, etc.

To install this "core" package, simply type the following from the R command prompt:
\begin{verbatim}
install.packages(
                 "Zelig",
                 repos = "http://r.iq.harvard.edu",
                 type  = "source"
                 )
\end{verbatim}

\subsection{Install Add-on Packages}

In addition to Zelig's core package, which exclusively contains simple regression models and a Developers' API for producing novel R packages, a myriad of Zelig add-on packages are available. These packages supplement Zelig's features, and add specialized, advanced models to Zelig.

\subsubsection{List of Available Packages}
These add-on packages include:

\begin{itemize}

	\item {\tt bivariate.zelig}: Bivariate Models for Logit and Probit Regressions

	\item {\tt gam.zelig}: Generalized Additive Models for Logit, Gaussian, Poisson and Probit Regressions

	\item {\tt gee.zelig}: Generalized Estimating Equation Models for Gamma, Logit, Gaussian, Poisson and Probit Regressions
	
	\item {\tt mixed.zelig}: Mixed Effect Models (Multi-level) for Gamma, Logit, Least-squares, Poisson and Probit Regressions
	
	\item {\tt multinomial.zelig}: Multinomial Models Logit and Probit Regressions
	
	\item {\tt ordinal.zelig}: Ordinal Models for Logit and Probit Regressions

	\item {\tt survey.zelig}: Survey-weighted Models for Gamma, Logit, Normal, Poisson and Probit Regressions

\end{itemize}

\subsubsection{Using {\tt install.packages}}
\begin{minipage}{\linewidth}
To download any of these packages independently, simply type the following from an R command prompt:
\begin{verbatim}
install.packages(
                 "MODEL NAME",
                 repos = "http://r.iq.harvard.edu/",
                 type  = "source"
                 )
\end{verbatim}

Where {\tt "MODEL NAME"} is replaced with the title of the Add-on packages in the above itemized list. For example, to download "Generalized Estimating Equation Models...", note that its package name is {\tt gee.zelig}, and type from the R command prompt:
\begin{verbatim}
install.packages(
                 "gee.zelig",
                 repos = "http://r.iq.harvard.edu/",
                 type  = "source"
                 )
\end{verbatim}
\end{minipage}

% Archived Installation: How to install Zelig via tarballs, etc.
%%\section{Install from Archived Zelig File}
%
%In addition to the interactive installation procedures - sections 3 and 4 - Zelig can be %installed 



\section{Post-Installation}

Barring any installation errors, Zelig and any add-on packages that were manually installed, should now be available from an R-session. Simply type from an R command prompt:

\begin{verbatim}
library(Zelig)
?Zelig
\end{verbatim}

To begin interacting and using the Zelig software package. Additionally, demo files can be listed via the command:

\begin{verbatim}
demo(package="Zelig")
\end{verbatim}








